# Reproducible Research: Peer Assessment 1

```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
```

## Loading and preprocessing the data
##### 1. Load the data (i.e. read.csv())
```{r, echo=TRUE, results='markup', warning=TRUE, message=TRUE}
activityData <- read.csv('activity.csv')
```
##### 2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r, echo=TRUE}
#activityData$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activityData$interval), format='%H:%M')
```

-----

## What is mean total number of steps taken per day?
##### 1. Calculate the total number of steps taken per day
```{r, echo=TRUE}
stepsByDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
stepsByDay
```

##### 2. Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
hist(stepsByDay, 
     breaks=seq(from=0, to=25000, by=2500),
     col="red", 
     xlab="Total steps per day", 
     ylim=c(0, 20), 
     main="Total number of steps taken per day")
```

##### 3. Calculate and report the mean and median total number of steps taken per day
```{r, echo=TRUE}
stepsByDayMean <- mean(stepsByDay)
stepsByDayMedian <- median(stepsByDay)
```
* Mean: `r stepsByDayMean`
* Median:  `r stepsByDayMedian`

-----

## What is the average daily activity pattern?
```{r, echo=TRUE}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)
```

##### 1. Make a time series plot
```{r, echo=TRUE}
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-Minute Interval") +
    ylab("Average Number of Steps") 
```

##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=TRUE}
mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
```

* Most Steps at: `r timeMostSteps`

----

## Imputing missing values
##### 1. Calculate and report the total number of missing values in the dataset 
```{r, echo=TRUE}
numMissingValues <- length(which(is.na(activityData$steps)))
```

* Number of missing values: `r numMissingValues`

##### 2. Devise a strategy for filling in all of the missing values in the dataset.
##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo=TRUE}
activityData <- activityData
activityData$steps <- impute(activityData$steps, fun=mean)
```


##### 4a. Make a histogram of the total number of steps taken each day 
```{r, echo=TRUE}
stepsByDay <- tapply(activityData$steps, activityData$date, sum)
hist(stepsByDay, 
     breaks=seq(from=0, to=25000, by=2500),
     col="red", 
     xlab="Total steps per day", 
     ylim=c(0, 30), 
     main="Total number of steps taken per day")
```

##### 4b. Calculate and report the mean and median total number of steps taken per day. 
```{r, echo=TRUE}
stepsByDayMean <- mean(stepsByDay)
stepsByDayMedian <- median(stepsByDay)
```
* Mean: `r stepsByDayMean`
* Median:  `r stepsByDayMedian`

----

## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r, echo=TRUE}
activityData$dateType <-  ifelse(as.POSIXlt(activityData$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

##### 2. Make a panel plot containing a time series plot

```{r, echo=TRUE}
averagedActivityData <- aggregate(steps ~ interval + dateType, data=activityData, mean)
ggplot(averagedActivityData, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-Minute Interval") + 
    ylab("Average Number of Steps")
```